PyTracer: Automatically Profiling Numerical Instabilities in Python

### 2022-05-13 IEEEtran/paper.tex
Fixed makefile
Fixed section 2 intro

\TG{I'm still unsure what to do with this part, it looks out of place.}
% \add{PyTracer follows a three steps approach:}
% \begin{enumerate}
%     \item \add{Automatic instrumentation of Python programs to save traces containing input and output values of all functions.}
%     \item \add{Traces aggregation to produce a statistics summary.}
%     \item \add{Interactive visualization to navigate through code execution.}
% \end{enumerate}

\pytracer adopts the three-step workflow shown in Figure~\ref{fig:workflow}.
First, \pytracer uses \textit{monkey patching} to replace the functions called
by the application with a wrapper that saves their arguments and returned values
in a trace file.
Second, it executes the application multiple times with noise injection, and
aggregates the resulting trace files to compute summary statistics about each
traced variable. Finally, a Plotly dashboard provides interactive visualizations
highlighting unstable code sections.

\add{
    The main goal of PyTracer is to help developpers visualize the numerical
    stability of Python applications. PyTracer saves time for the user by
    instrumenting and aggregating values. In the same way that GDB makes the user
    save time when debugging an application by providing all information in one
    layout, PyTracer allows the user to visualize the overall numerical quality to
    quickly detect instabilities. The PyTracer visualization layout improves result
    analysis in any case since 1) the user does not have to manually instrument its
    code to dump raw data and creates specific post-processing scripts to visualize
    it 2) PyTracer allows the user to export the raw data traced to extend the
    analysis. 
}

\add{
    The main challenge addressed by PyTracer is the transparent instrumentation of Python 
    applications. Indeed, since Python is an interpreted language, one has no complete 
    view of the function that will be executed by a given program. 
    Moreover, for Python applications used in HPC environments, one has
    most of the time no access to source codes or with limited write access.
    Therefore, PyTracer tackles this issue by instrumenting the function dynamically 
    thanks to the Python introspection.
}

\add{
    The trace-based approach used by PyTracer permits to limit the 
    overhead at runtime since the statistics computation is reported to the postprocess analysis.
    However, it limits the aggregation to deterministic executions,
    although control-flow analysis can help to merge divergent execution paths.
    Hence, PyTracer is restricted to single-threaded applications even if its usage
    is limited in Python due to the Global Interperter Lock (GIL) mechanism
    that allows only one thread to hold the control of the Python interpreter.
}

>>>>

% \add{PyTracer follows a three steps approach:}
% \begin{enumerate}
%     \item \add{Automatic instrumentation of Python programs to save traces containing input and output values of all functions.}
%     \item \add{Traces aggregation to produce a statistics summary.}
%     \item \add{Interactive visualization to navigate through code execution.}
% \end{enumerate}

\add{To detect numerical instabilities, \pytracer works in three main steps:}
\begin{enumerate}
\item It instruments Python programs and their module dependencies automatically, to save the input and output values of functions in a trace file.
\item It executes the instrumented program several times using Monte-Carlo Arithmetic to inject perturbations in function outputs.
\item It aggregates all trace files to compute statistics about the traced values, and visualize them in a GUI.
\end{enumerate}
\add{
    To implement these steps, \pytracer adopts the workflow shown in Figure~\ref{fig:workflow}.
At step 1, \pytracer uses \textit{monkey patching} to replace the functions called
by the application with a wrapper that saves their arguments and returned values
in a trace file. At step 2, it executes the application multiple times with Monte-Carlo Arithmetic as implemented in Fuzzy, a collection 
of scientific computing libraries recompiled with the Verificarlo tool. It then
aggregates the resulting trace files to compute summary statistics on
traced variable. At step 3, a Plotly dashboard provides interactive visualizations
highlighting unstable code sections.
}

% \add{
%     The main goal of PyTracer is to help developpers visualize the numerical
%     stability of Python applications. PyTracer saves time for the user by
%     instrumenting and aggregating values. In the same way that GDB makes the user
%     save time when debugging an application by providing all information in one
%     layout, PyTracer allows the user to visualize the overall numerical quality to
%     quickly detect instabilities. The PyTracer visualization layout improves result
%     analysis in any case since 1) the user does not have to manually instrument its
%     code to dump raw data and creates specific post-processing scripts to visualize
%     it 2) PyTracer allows the user to export the raw data traced to extend the
%     analysis. 
% }

% \add{
%     The main challenge addressed by PyTracer is the transparent instrumentation of Python 
%     applications. Indeed, since Python is an interpreted language, one has no complete 
%     view of the function that will be executed by a given program. 
%     Moreover, for Python applications used in HPC environments, one has
%     most of the time no access to source codes or with limited write access.
%     Therefore, PyTracer tackles this issue by instrumenting the function dynamically 
%     thanks to the Python introspection.
% }

% \add{
%     The trace-based approach used by PyTracer permits to limit the 
%     overhead at runtime since the statistics computation is reported to the postprocess analysis.
%     However, it limits the aggregation to deterministic executions,
%     although control-flow analysis can help to merge divergent execution paths.
%     Hence, PyTracer is restricted to single-threaded applications even if its usage
%     is limited in Python due to the Global Interperter Lock (GIL) mechanism
%     that allows only one thread to hold the control of the Python interpreter.
% }
